The impact of cosmic variance on simulating weak lensing surveys

### 2014-12-01 Writeup/paper_v3.tex
comment on current surveys as well in section on impact

One method that is commonly used to test for the presence of
systematic errors in the shear estimation process is image simulation,
where we can cleanly test whether our methods of shear estimation
recover the ground truth. This is a valuable test, considering the
numerous sources of additive and multiplicative bias such as a
mismatch between galaxy model assumptions and actual galaxy light
profiles \citep[e.g.,][]{2010MNRAS.404..458V, 2010A&A...510A..75M}, biases due to the effects of pixel noise on the
shear estimates
\citep{2012MNRAS.427.2711K,2012MNRAS.424.2757M,2012MNRAS.425.1951R},
and ellipticity gradients \citep{2010MNRAS.406.2793B}.  These biases
often differ for galaxies with different morphologies (e.g., disks
vs. ellipticals), sizes, $S/N$, and shape \citep{2010MNRAS.405.2044B,2012MNRAS.423.3163K}. 
A general requirement for simulations used to test shear recovery is
that they should be as realistic as possible.

@@

% \rachel{As we discussed, you should get rid of the early parts of this
%   regarding second moments, and replace it with a brief description of
%   what shape estimates we use from GalSim (re-Gaussianization).}
% 
% {\bf Should explain in brief Claire's work.} \\
\begin{figure*}
 \centering
 \includegraphics[width=2.2\columnwidth]{sample_histograms}
 \caption{The distributions of the morphological parameters of interest, namely the distortion (top left, from \sersic fits, and top right, from re-Gaussianization), \sersic index (bottom left) and bulge-to-total ratio (bottom right) are presented. 
          The shapes of these distributions depend on the way the volume-limiting process is carried out. \s$0$ refers to the flux-limited ($m_\text{F814}\le23.5$) sample
          and \s$1$, \s$2$ and \s$3$ refer to the volume-limited
          samples, as discussed in
          Sec.~\ref{sub:volumelimiting}. 
%           \rachel{I am confused by the
%             top line: first, distortion is $(1-q^2)/(1+q^2)$, not
%             $(1-q)/(1+q)$.  Second, I am not sure which shapes you are
%           plotting - from Claire or from re-Gaussianization.  It seems
%         redundant to have two plots of the same thing, so I hope that
%         one is from Claire and one from re-Gaussianization, but these
%         needs to be indicated clearly.  Finally, we have the option of
%       deliberately plotting different things (like $q$ from Claire
%       vs. distortion from re-Gaussianization) or explicitly choosing
%       to plot the same thing (like total distortion from Claire
%       vs. total distortion from re-Gaussianization).  I'm not sure
%       which is best; the latter would allow a comparison between the
%       two types of shape estimates, which could be interesting, so I'm
%       inclined to say you should do that.}} \arun{Done.}
  }
  \label{fig:sample_histograms}
\end{figure*}

@@

\begin{table}
 \centering
 \begin{tabular}[\columnwidth]{ | l | c | c | c | }
  \hline
  %\parbox[t]{0.15\columnwidth}{Redshift bins} & \parbox[t]{0.15\columnwidth}{Tests} & \parbox[t]{0.15\columnwidth}{Luminosity-selected} & \parbox[t]{0.15\columnwidth}{Luminosity-selected with B-band evolution} & \parbox[t]{0.15\columnwidth}{Stellar-mass-selected} \\
  Redshift bins & \s$1$ & \s$2$ & \s$3$ \\
  \hline
  All overdense vs.  & \scinot{1.1}{-4} & \scinot{2.6}{-5} & \scinot{1.9}{-6} \\
  All underdense     & \scinot{1}{-5} & $<\scinot{1}{-5}$ & $<\scinot{1}{-5}$ \\ \hline 
  $[0.65,0.75]$ (OD) vs.$\!\!\!$ & 0.61 & 0.43 & 0.23 \\
  $[0.80,0.85]$ (OD) & 0.49 & 0.24 & 0.13 \\ \hline
  $[0.65,0.75]$ (OD) vs.$\!\!\!$ & \scinot{5.8}{-4} & \scinot{1.5}{-5} & \scinot{3.5}{-6} \\
  $[0.55,0.65]$ (UD) & \scinot{9.8}{-4} & $<\scinot{1}{-5}$ & $<\scinot{1}{-5}$ \\ \hline
 \end{tabular}
 \caption{$p$-values from the Kolmogorov-Smirnov (top) and
   Anderson-Darling (bottom) tests obtained by comparing the
   distributions of axis ratios for three cases: \emph{all} overdense
   (OD) vs. \emph{all} underdense (UD), two overdense bins that are
   not very separated in redshift, and a pair of adjacent overdense
   and underdense bins. \s$1$, \s$2$, \s$3$ refer to the three
   different types of volume-limited samples.
          The Anderson-Darling $p$-values are computed only up to 5
          decimal places, so values that were given as zero are
          denoted $<\scinot{1}{-5}$.
% \rachel{What are the 0 entries in this and other tables?  Just how close to zero are they?  Since 0 is not really achieve, I recommend defining a value (such as $10^{-10}$ below which you will just say $<10^{-10}$ and give the value if it's above that, even if it seems really small to you, like $10^{-8}$.  Also, say $1.1\times 10^{-4}$ rather than $1.1e-4$.}
          }
 \label{table:pvalues_all}
\end{table}

\begin{table}
 \centering
 \begin{tabular}[\columnwidth]{ | l | c | c | c | }
  \hline
  %\parbox[t]{0.15\columnwidth}{Redshift bins} & \parbox[t]{0.15\columnwidth}{Tests} & \parbox[t]{0.15\columnwidth}{Luminosity-selected} & \parbox[t]{0.15\columnwidth}{Luminosity-selected with B-band evolution} & \parbox[t]{0.15\columnwidth}{Stellar-mass-selected} \\
  Redshift bins & \s$1$ & \s$2$ & \s$3$ \\
  \hline
  All overdense vs. & \scinot{5.6}{-4} & \scinot{1.0}{-4} & \scinot{3.3}{-6} \\
  All underdense    & \scinot{3}{-5} & \scinot{1}{-5} & $<\scinot{1}{-5}$ \\ \hline 
  $[0.65,0.75]$ (OD) vs.$\!\!\!$ & 0.96 & 0.75 & 0.54 \\
  $[0.80,0.85]$ (OD) & 0.52 & 0.34 & 0.23 \\ \hline
  $[0.65,0.75]$ (OD) vs.$\!\!\!$ & \scinot{6.0}{-3} & \scinot{2.5}{-4} & \scinot{2.4}{-4} \\
  $[0.55,0.65]$ (UD) & \scinot{1.2}{-2} & \scinot{2.5}{-4} & \scinot{5}{-5} \\ \hline
 \end{tabular}
 \caption{$p$-values from the Kolmogorov-Smirnov (top) and
   Anderson-Darling (bottom) obtained by comparing the second
   moments-based distortion for the same three cases as in
   Table~\ref{table:pvalues_all}. 
           The Anderson-Darling $p$-values are computed only up to 5 decimal places. }
 \label{table:pvalues_momentbased_all}
\end{table}

@@

\rachel{As mentioned previously, {\em HST} data will be used by future weak
lensing surveys to characterize the galaxy population in several
ways.  In this section, we estimate how the results shown in
Sec.~\ref{S:results} can affect estimates of shear calibration when
using {\em HST} to characterize the galaxy population.  We also
discuss the situations in which this is likely to be important for
current and future surveys from the space and ground.}

@@

\subsection{\rachel{Effective impact on future surveys}}

\rachel{In addition to the order of magnitude of these effects
  presented in the previous subsection, it is important to bear in
  mind how future surveys plan to use {\em HST} data.}

@@

\rachel{In contrast, for current and future ground-based surveys such
  as LSST, high-resolution space data will play a more important role
  in the understanding of the galaxy 
  intrinsic ellipticity and morphology distributions, since even a deep
  field in a ground-based survey faces fundamental resolution limits
  that prevent the derivation of detailed information about the faint
  galaxy population.  Galaxies that are near the resolution limit for
  a ground-based survey are still very well-resolved in {\em HST},
  making it the best resource for detailed information about them.
  Once the Euclid deep field is publicly available, the information
  from it will be beneficial to ground-based surveys as well.}

@@

In this study, we have shown that the shape distributions of galaxies
(to a statistically significant degree) and morphological parameters
like \sersic $n$ and bulge-to-total ratios (more marginally) depend on
the local environments when dividing up the COSMOS sample into
redshift slices along the line of sight.  The redshift slices used for
our primary analysis had
a width of $\Delta z=0.05$.  Our findings are robust to the choice of
shape estimator from \sersic profile fits vs. using centrally weighted
moments-based shear estimates.

>>>>

One method that is commonly used to test for the presence of
systematic errors in the shear estimation process is image simulation,
where we can cleanly test whether our methods of shear estimation
recover the ground truth. This is a valuable test, considering the
numerous sources of additive and multiplicative bias such as a
mismatch between galaxy model assumptions and actual galaxy light
profiles \citep[e.g.,][]{2010MNRAS.404..458V, 2010A&A...510A..75M}, biases due to the effects of pixel noise on the
shear estimates
\citep{2012MNRAS.427.2711K,2012MNRAS.424.2757M,2012MNRAS.425.1951R},
and ellipticity gradients \citep{2010MNRAS.406.2793B}.  These biases
often differ for galaxies with different morphologies (e.g., disks
vs.\ ellipticals), sizes, $S/N$, and shape \citep{2010MNRAS.405.2044B,2012MNRAS.423.3163K}. 
A general requirement for simulations used to test shear recovery is
that they should be as realistic as possible.

@@

% \rachel{As we discussed, you should get rid of the early parts of this
%   regarding second moments, and replace it with a brief description of
%   what shape estimates we use from GalSim (re-Gaussianization).}
% 
% {\bf Should explain in brief Claire's work.} \\
\begin{figure*}
 \centering
 \includegraphics[width=2.2\columnwidth]{sample_histograms}
 \caption{The distributions of the morphological parameters of interest, namely the distortion (top left, from \sersic fits, and top right, from re-Gaussianization), \sersic index (bottom left) and bulge-to-total ratio (bottom right) are presented. 
          The shapes of these distributions depend on the way the volume-limiting process is carried out. \s$0$ refers to the flux-limited ($m_\text{F814}\le23.5$) sample
          and \s$1$, \s$2$ and \s$3$ refer to the volume-limited
          samples, as discussed in
          Sec.~\ref{sub:volumelimiting}. 
%           \rachel{I am confused by the
%             top line: first, distortion is $(1-q^2)/(1+q^2)$, not
%             $(1-q)/(1+q)$.  Second, I am not sure which shapes you are
%           plotting - from Claire or from re-Gaussianization.  It seems
%         redundant to have two plots of the same thing, so I hope that
%         one is from Claire and one from re-Gaussianization, but these
%         needs to be indicated clearly.  Finally, we have the option of
%       deliberately plotting different things (like $q$ from Claire
%       vs.\ distortion from re-Gaussianization) or explicitly choosing
%       to plot the same thing (like total distortion from Claire
%       vs.\ total distortion from re-Gaussianization).  I'm not sure
%       which is best; the latter would allow a comparison between the
%       two types of shape estimates, which could be interesting, so I'm
%       inclined to say you should do that.}} \arun{Done.}
  }
  \label{fig:sample_histograms}
\end{figure*}

@@

\begin{table}
 \centering
 \begin{tabular}[\columnwidth]{ | l | c | c | c | }
  \hline
  %\parbox[t]{0.15\columnwidth}{Redshift bins} & \parbox[t]{0.15\columnwidth}{Tests} & \parbox[t]{0.15\columnwidth}{Luminosity-selected} & \parbox[t]{0.15\columnwidth}{Luminosity-selected with B-band evolution} & \parbox[t]{0.15\columnwidth}{Stellar-mass-selected} \\
  Redshift bins & \s$1$ & \s$2$ & \s$3$ \\
  \hline
  All overdense vs.\  & \scinot{1.1}{-4} & \scinot{2.6}{-5} & \scinot{1.9}{-6} \\
  All underdense     & \scinot{1}{-5} & $<\scinot{1}{-5}$ & $<\scinot{1}{-5}$ \\ \hline 
  $[0.65,0.75]$ (OD) vs.$\!\!\!$ & 0.61 & 0.43 & 0.23 \\
  $[0.80,0.85]$ (OD) & 0.49 & 0.24 & 0.13 \\ \hline
  $[0.65,0.75]$ (OD) vs.$\!\!\!$ & \scinot{5.8}{-4} & \scinot{1.5}{-5} & \scinot{3.5}{-6} \\
  $[0.55,0.65]$ (UD) & \scinot{9.8}{-4} & $<\scinot{1}{-5}$ & $<\scinot{1}{-5}$ \\ \hline
 \end{tabular}
 \caption{$p$-values from the Kolmogorov-Smirnov (top) and
   Anderson-Darling (bottom) tests obtained by comparing the
   distributions of axis ratios for three cases: \emph{all} overdense
   (OD) vs.\ \emph{all} underdense (UD), two overdense bins that are
   not very separated in redshift, and a pair of adjacent overdense
   and underdense bins. \s$1$, \s$2$, \s$3$ refer to the three
   different types of volume-limited samples.
          The Anderson-Darling $p$-values are computed only up to 5
          decimal places, so values that were given as zero are
          denoted $<\scinot{1}{-5}$.
% \rachel{What are the 0 entries in this and other tables?  Just how close to zero are they?  Since 0 is not really achieve, I recommend defining a value (such as $10^{-10}$ below which you will just say $<10^{-10}$ and give the value if it's above that, even if it seems really small to you, like $10^{-8}$.  Also, say $1.1\times 10^{-4}$ rather than $1.1e-4$.}
          }
 \label{table:pvalues_all}
\end{table}

\begin{table}
 \centering
 \begin{tabular}[\columnwidth]{ | l | c | c | c | }
  \hline
  %\parbox[t]{0.15\columnwidth}{Redshift bins} & \parbox[t]{0.15\columnwidth}{Tests} & \parbox[t]{0.15\columnwidth}{Luminosity-selected} & \parbox[t]{0.15\columnwidth}{Luminosity-selected with B-band evolution} & \parbox[t]{0.15\columnwidth}{Stellar-mass-selected} \\
  Redshift bins & \s$1$ & \s$2$ & \s$3$ \\
  \hline
  All overdense vs.\ & \scinot{5.6}{-4} & \scinot{1.0}{-4} & \scinot{3.3}{-6} \\
  All underdense    & \scinot{3}{-5} & \scinot{1}{-5} & $<\scinot{1}{-5}$ \\ \hline 
  $[0.65,0.75]$ (OD) vs.$\!\!\!$ & 0.96 & 0.75 & 0.54 \\
  $[0.80,0.85]$ (OD) & 0.52 & 0.34 & 0.23 \\ \hline
  $[0.65,0.75]$ (OD) vs.$\!\!\!$ & \scinot{6.0}{-3} & \scinot{2.5}{-4} & \scinot{2.4}{-4} \\
  $[0.55,0.65]$ (UD) & \scinot{1.2}{-2} & \scinot{2.5}{-4} & \scinot{5}{-5} \\ \hline
 \end{tabular}
 \caption{$p$-values from the Kolmogorov-Smirnov (top) and
   Anderson-Darling (bottom) obtained by comparing the second
   moments-based distortion for the same three cases as in
   Table~\ref{table:pvalues_all}. 
           The Anderson-Darling $p$-values are computed only up to 5 decimal places. }
 \label{table:pvalues_momentbased_all}
\end{table}

@@

\rachel{As mentioned previously, {\em HST} data will be used by
  current and future weak
lensing surveys to characterize the galaxy population in several
ways.  In this section, we estimate how the results shown in
Sec.~\ref{S:results} can affect estimates of shear calibration when
using {\em HST} to characterize the galaxy population.  We also
discuss the situations in which this is likely to be important for
current and future surveys from the space and ground.}

@@

\subsection{\rachel{Effective impact on current and future surveys}}

\rachel{In addition to the order of magnitude of these effects
  presented in the previous subsection, it is important to bear in
  mind how current and future surveys plan to use {\em HST} data.}

@@

\rachel{In contrast, for ground-based surveys such
  as LSST, high-resolution space data will play a more important role
  in the understanding of the galaxy 
  intrinsic ellipticity and morphology distributions, since even a deep
  field in a ground-based survey faces fundamental resolution limits
  that prevent the derivation of detailed information about the faint
  galaxy population.  Galaxies that are near the resolution limit for
  a ground-based survey are still very well-resolved in {\em HST},
  making it the best resource for detailed information about them.
  Once the Euclid deep field is publicly available, the information
  from it will be beneficial to ground-based surveys as well.}

\rachel{Finally, our results for the effect on the intrinsic
  ellipticity distributions suggest that even current surveys should
  be careful to avoid this effect.  While use of COSMOS without
  accounting for the effect that we have identified will cause a bias
  that is similar to the final requirements on the shear systematic
  errors for Stage III surveys, there are other elements in the
  systematic error budget, so ideally this effect should be mitigated
  somewhat using, for example, the wider redshift binning strategy
  that we tested in Sec.~\ref{sec:mitigation}.  Use of significantly smaller fields
  than COSMOS, while possibly helpful in building up a deeper galaxy
  sample, should naturally increase the impact of cosmic
  variance on the training dataset, which should be quantified as part
  of the systematic error budget.}

@@

In this study, we have shown that the shape distributions of galaxies
(to a statistically significant degree) and morphological parameters
like \sersic $n$ and bulge-to-total ratios (more marginally) depend on
the local environments when dividing up the COSMOS sample into
redshift slices along the line of sight.  The redshift slices used for
our primary analysis had
a width of $\Delta z=0.05$.  Our findings are robust to the choice of
shape estimator from \sersic profile fits vs.\ using centrally weighted
moments-based shear estimates.
